---
title: "Notification Strategies for Tax Recovery"
author: "Michael Chirico^[The research presented here was supported in part by the Institute of Education Sciences, U.S. Department of Education, through Grant \\#R305B090015 to the University of Pennsylvania. The opinions expressed are those of the presenter and do not represent the views of the Institute or the U.S. Department of Education.], Robert Inman, Charles Loeffler, John MacDonald, Holger Sieg"
date: "March 1, 2016"
output: 
  beamer_presentation: 
    keep_tex: yes
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(cache=TRUE)
knitr::read_chunk("round_2_analysis.R")
```


## Motivation
```{r first, echo=FALSE, results = "hide"}
<<setup>>
  x <- dol.form(properties[,sum(total_due)])
```

I have access to `r x`.



This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

[^1]: The research presented here was supported in part by the Institute of Education Sciences, U.S. Department of Education, through Grant \#R305B090015 to the University of Pennsylvania. The opinions expressed are those of the presenter and do not represent the views of the Institute or the U.S. Department of Education.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Code and Output

```{r}
summary(cars)
```

## Slide with Plot

```{r, echo=FALSE}
plot(cars)
```

