---
title: "Notification Strategies for Tax Recovery"
author: "Michael Chirico^[The research presented here was supported in part by the Institute of Education Sciences, U.S. Department of Education, through Grant \\#R305B090015 to the University of Pennsylvania. The opinions expressed are those of the presenter and do not represent the views of the Institute or the U.S. Department of Education.], Robert Inman, Charles Loeffler, John MacDonald, Holger Sieg"
date: "March 1, 2016"
output: 
  beamer_presentation: 
    keep_tex: yes
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(cache=TRUE)
knitr::read_chunk("round_2_analysis.R")
```


## Motivation
```{r first, echo=FALSE, results = "hide"}
<<setup>>
library(data.table)
library(funchir)
out_taxes <- fread(wds["data"] %+% "Real_Estate_Tax_Balances_151222.csv",
                   select = c("Tax Period", "Total")
                   )[`Tax Period` != 2016]
out_taxes[ , due := as.numeric(gsub("\\$", "", Total))]

taxes_owed <- out_taxes[ , dol.form(sum(due))]

jul_del_lien_cont_ep <-
  owners[ , to.pct(mean(ever_paid_jul[treat7 == "Lien"]) - 
                     mean(ever_paid_jul[treat7 == "Control"]), dig = 0)]

dec_del_lien_cont_ep <-
  owners[ , to.pct(mean(ever_paid_dec[treat7 == "Lien"]) -
                     mean(ever_paid_dec[treat7 == "Control"]), dig = 0)]
```

* Tax non-compliance is ubiquitous in both developed and undeveloped economies

    + OECD estimates: 14.2% in developed economies, 37% in undeveloped

* The City of Philadelphia is currently owed `r taxes_owed` in outstanding Real Estate Taxes.

    + About 15% of all properties in the city were out of compliance in March 2015.
   
    + Roughly 5% of the City's annual operating budget.

## Motivation

* Much existing literature focuses on sleight-of-hand -- non- or under-reporting of assets/income in environments where enforcement is scant or verifiability is costly.

* No such fallback for real estate taxes
 
    + Costs of establishing an office of assessment are sunk
   
    + Exact tax value of each property is determined by the city and reported to the taxpayer (reversal)
    
* Alternative explanations

    + Discontent with provided public goods
    
    + Low tax morale
    
    + Missing social norms
    
    + Perceived unfairness of taxation

## Motivation

* Costs of containing culture of non-payment can quickly spiral

    + Fewer people pay $\Rightarrow$ harder to exact widespread compliance from a fixed budget $\Rightarrow$ spread of noncompliance as perceived likelihood of apprehension decreases
    
    + Cost of enforcement through augmented seizures can soon exceed potential recoveries
    
    + To sustain public goods outlay, tax rates must rise -- potentially inducing further noncompliance
    
    + Alternative of reduced expenditures may also create rogues through diminished tax morale
    
* Low-cost interventions offer a potential lighthouse in this hurricane spiral

    + Targeted messaging in mailers
    
    + Notification frequency
    
    + Shape, color, style manipulation
    
    + Public shaming
    
## Overview of Treatments

* Simultaneous (randomized) evaluation and head-to-head comparison of a variety of letter-based recovery nudges

    + Message targeting: 3 pairs of treatments with wording varied to target different compliance motives
    
    + Style cues: Half of treatments were delivered in large-format ($9$'' x $12$'')envelopes, the rest in standard-sized ($4 \frac18$'' x $9 \frac12$'') envelopes.
    
    + Frequency: Some owners with multiple properties only received treatments on a subset of their holdings.
    
* All told, 15 treatment groups (including Holdout & Control groups) on `r prettyNum(nrow(properties), big.mark = ",")` properties with `r prettyNum(nrow(owners), big.mark = ",")` owners.
    
## Overview of Results

* Evidence of strong initial effect for deterrence-oriented treatments

    + Roughly `r jul_del_lien_cont_ep`% higher participation (any payment) rates for owners in the Lien threat treament after one month.
    
    + Also evidence of long-run effect mediation -- by the end of 2015 (six months post-treatment), the advantage had shrunk to roughly `r dec_del_lien_cont_ep`%.

* No evidence whatsoever that letter size affects compliance rates

* Support for hypothesis that more letters lead to more participation

## Experimental Framework



## Slide with R Code and Output

```{r}
summary(cars)
```

## Slide with Plot

```{r, echo=FALSE}
plot(cars)
```

