---
title: "Notification Strategies for Tax Recovery"
author: "Michael Chirico^[The research presented here was supported in part by the Institute of Education Sciences, U.S. Department of Education, through Grant \\#R305B090015 to the University of Pennsylvania. The opinions expressed are those of the presenter and do not represent the views of the Institute or the U.S. Department of Education.], Robert Inman, Charles Loeffler, John MacDonald, Holger Sieg"
date: "March 1, 2016"
output: 
  beamer_presentation: 
    keep_tex: yes
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(cache=TRUE)
knitr::read_chunk("round_2_analysis.R")
```


## Motivation
```{r first, echo=FALSE, results = "hide"}
<<setup>>
library(funchir)
out_taxes <- fread(wds["data"]%+%"Real_Estate_Tax_Balances_151222.csv",
                   select=c("Tax Period","Total")
                   )[`Tax Period`!=2016]
out_taxes[ , due := as.numeric(gsub("\\$","",Total))]

taxes_owed <- out_taxes[,dol.form(sum(due))]
```


* Tax non-compliance is ubiquitous in both developed and undeveloped economies

    + OECD estimates: 14.2% in developed economies, 37% in undeveloped

* The City of Philadelphia is currently owed `r taxes_owed` in outstanding Real Estate Taxes.

    + About 15% of all properties in the city were out of compliance in March 2015.
   
    + Roughly 5% of the City's annual operating budget.

## Motivation

* Much existing literature focuses on sleight-of-hand -- non- or under-reporting of assets/income in environments where enforcement is scant or verifiability is costly.

* No such fallback for real estate taxes
 
    + Costs of establishing an office of assessment are sunk
   
    + Exact tax value of each property is determined by the city and reported to the taxpayer (reversal)

## Slide with R Code and Output

```{r}
summary(cars)
```

## Slide with Plot

```{r, echo=FALSE}
plot(cars)
```

